---
title: "2-combined-analysis"
author: "IndoorOutdoor Team""
date: "6/08/2017"
output: html_document
---

## Loading BetyDB into environment and joining to make main table

```{r db-connection}
library(dplyr)

bety_src <- src_postgres(dbname = "bety", 
                password = 'bety', 
                host = 'terra-bety.default', 
                user = 'bety', 
                port = 5432)

# to see all available columns in traits table
original_traits <- tbl(bety_src, 'traits') %>%
  collect(n=1)

# local version of variables for reference
variables_local <- tbl(bety_src, 'variables', n = Inf) %>%
  mutate(variable_id = id, variable_name = name) %>%
  dplyr::select(variable_id, variable_name, description, units) %>%
  collect()

traits <- tbl(bety_src, 'traits', n = Inf) %>%
  mutate(trait_id = id) %>%
  dplyr::select(trait_id, site_id, specie_id, cultivar_id, date, mean, variable_id, method_id, treatment_id, entity_id)

variables <- tbl(bety_src, 'variables', n = Inf) %>%
  mutate(variable_id = id, variable_name = name) %>%
  dplyr::select(variable_id, variable_name)

cultivars <- tbl(bety_src, 'cultivars', n = Inf) %>%
  mutate(cultivar_id = id, cultivar = name) %>%
  dplyr::select(cultivar_id, cultivar)

entities <- tbl(bety_src, 'entities', n = Inf) %>%
  mutate(entity_name = name, entity_id = id) %>%
  dplyr::select(entity_name, entity_id)

sites <- tbl(bety_src, 'sites', n = Inf) %>%
  mutate(site_id = id) %>%
  dplyr::select(site_id, city, state, country, notes, sitename)

treatments <- tbl(bety_src, 'treatments', n = Inf) %>%
  mutate(treatment_id = id, treatment_definition = definition, treatment_name = name) %>%
  dplyr::select(treatment_id, treatment_name, treatment_definition) 

# looking for when each season began
experiments <- tbl(bety_src, 'experiments', n = Inf) %>%
  #mutate(treatment_id = id, treatment_definition = definition, treatment_name = name) %>%
  #dplyr::select(treatment_id, treatment_name, treatment_definition) 
  collect(n = Inf)

# join relevant tables together
joined_table <- traits %>%
  left_join(variables, by = 'variable_id') %>%
  left_join(cultivars, by = 'cultivar_id') %>%
  left_join(entities, by = 'entity_id') %>%
  left_join(sites, by = 'site_id') %>%
  left_join(treatments, by = 'treatment_id') %>%
  dplyr::select(trait_id, date, mean, variable_name, sitename, treatment_name, cultivar)
```

## Filtering Table and Definining New Variables

```{r variable-tables}
library(lubridate)

# filter table for measurements that we care about
filtered_table <- filter(joined_table, variable_name %in% c("height", "canopy_cover", "canopy_height", "perimeter", "aboveground_dry_biomass", "leaf_length", "leaf_width", "plant_height", "aboveground_fresh_biomass", "growth_respiration_coefficient", "germination_score", "stem_diameter", "emergence_count", "cultivar"))

height_table <- filter(filtered_table, variable_name %in% c("canopy_height", "plant_height", "height")) %>%
  collect(n = Inf)

# filter table for rows without date
height_table <- filter(height_table, !is.na(height_table$date))

# define variable indoor_outdoor based on sitename
height_table$indoor_outdoor <- ifelse(height_table$sitename == 'Danforth Plant Science Center Bellweather Phenotyping Facility', 'Indoor', 'Outdoor')

# define variable height(cm)
# canopy_height is in , plant_height is in 
# range(filter(height_table, height_table$variable_name == "canopy_height")$mean) gives 5 and 270
# range(filter(height_table, height_table$variable_name == "plant_height")$mean) gives 17 and 1671, variables table tells us that the unit is cm but seems to be mm
height_table$height_cm <- if_else(height_table$variable_name == "canopy_height", height_table$mean, if_else(height_table$variable_name == "plant_height", height_table$mean/10, height_table$mean))

# define variable Date that is a date type
height_table$Date <- as.Date(substr(height_table$date, 0, 10), format = "%Y-%m-%d")

# define new variable season based on which season the data was taken
summer2014 = interval(as.Date("2014-05-01", format = "%Y-%m-%d"), as.Date("2014-07-31", format = "%Y-%m-%d"))
summer2016 = interval(as.Date("2016-05-01", format = "%Y-%m-%d"), as.Date("2016-07-31", format = "%Y-%m-%d"))
fall2016 = interval(as.Date("2016-08-15", format = "%Y-%m-%d"), as.Date("2016-12-31", format = "%Y-%m-%d"))

# checked whether all data in height_table fall into one of three seasons defined
height_table$season <- if_else(height_table$Date %within% summer2014, "Summer 2014", if_else(height_table$Date %within% summer2016, "Summer 2016", "Fall 2016"))

# define new variable age based on when the season began
# 2016 season start dates came from experiments table, 2014 season start date came from Dr. LeBauer
summer2014start = as.Date("2014-05-27", format = "%Y-%m-%d")
summer2016start = as.Date("2016-04-19", format = "%Y-%m-%d")
fall2016start = as.Date("2016-08-03", format = "%Y-%m-%d")

height_table$age <- if_else(height_table$season == "Summer 2014", as.numeric(height_table$Date - summer2014start), if_else(height_table$season == "Summer 2016", as.numeric(height_table$Date - summer2016start), as.numeric(height_table$Date - fall2016start)))

height_table$site_season <- paste(height_table$indoor_outdoor, " - ", height_table$season)

canopy_height_plant_height <- filter(height_table, variable_name %in% c("canopy_height", "plant_height")) %>%
  collect(n = Inf)

# look at which sites have which variables and cultivars
site_variable <- height_table %>%
  group_by(indoor_outdoor, variable_name, cultivar) %>%
  summarize(n = n()) %>%
  collect(n = Inf)

shared_cultivar <- filter(site_variable, cultivar %in% c('BTx642', 'PI_564163', 'Tx430', 'TX7000')) %>%
  collect(n = Inf)
```

## Plotting

```{r plotting-height-vs-age}
library(ggplot2)
library(gridExtra)

# plotting height against age
ggplot(canopy_height_plant_height, aes(x = age, y = height_cm)) +
  geom_point(aes(color = site_season, alpha = 1/10)) +
  ylab("Height (cm)") +
  ggtitle("Height")

ggplot(canopy_height_plant_height, aes(x = age, y = height_cm)) +
  geom_point(aes(color = site_season, alpha = 1/10)) +
  facet_wrap(~site_season) +
  ylab("Height (cm)") +
  ggtitle("Height by Season and Site")
  
# shared cultivars between indoor and outdoor

ggplot(filter(canopy_height_plant_height, cultivar %in% c('BTx642', 'PI_564163', 'Tx430', 'TX7000')), aes(x = age, y = height_cm)) +
  geom_point(aes(color = site_season, alpha = 1/10)) +
  facet_wrap(~cultivar) +
  ylab("Height (cm)") +
  ggtitle("Shared Cultivars")

# indoor plants
indoor_height <- filter(canopy_height_plant_height, indoor_outdoor == 'Indoor')

ggplot(indoor_height, aes(x = age, y = height_cm)) +
  geom_point(aes(color = cultivar, alpha = 1/10)) +
  facet_wrap(~cultivar) +
  ylab("Height (cm)") +
  ggtitle("Indoor Plants by Cultivar")

ggplot(indoor_height, aes(x = age, y = height_cm)) +
  geom_point(aes(color = treatment_name, alpha = 1/10)) +
  facet_wrap(~treatment_name) +
  ylab("Plant Height (Indoor)") +
  ggtitle("Indoor Plants by Treatment")

# outdoor plants
outdoor_height <- filter(canopy_height_plant_height, indoor_outdoor == 'Outdoor')

ggplot(outdoor_height, aes(x = age, y = height_cm)) +
  geom_point(aes(color = season, alpha = 1/10)) +
  ylab("Height (cm)") +
  ggtitle("Outdoor Plants by Season")

```

### Modeling - Emily

## Plots
```{r}
ggplot(indoor_height, aes(x = Date, y = height_cm)) +
  geom_point(alpha = 1/10) +
  ylab("Height (cm)") +
  ggtitle("Indoor Plants - Plant Height")

ggplot(filter(height_table, variable_name == "height"), aes(x = Date, y = height_cm)) +
  geom_point(alpha = 1/10) +
  ylab("Height (cm)") +
  ggtitle("Outdoor Plants - Height")

ggplot(filter(height_table, variable_name == "canopy_height"), aes(x = Date, y = height_cm)) +
  geom_point(alpha = 1/10) +
  ylab("Height (cm)") +
  ggtitle("Outdoor Plant - Canopy Height")

```

## Linear Regression!

```{r}
indoor_growthmodel <- lm(height_cm ~ age, data = indoor_height)
summary(indoor_growthmodel)

outdoorsummer_growthmodel <- lm(height_cm ~ age, data = outdoor_height[outdoor_height$season == "Summer 2016",])
summary(outdoorsummer_growthmodel)

outdoorfall_growthmodel <- lm(height_cm ~ age, data = outdoor_height[outdoor_height$season == "Fall 2016",])
summary(outdoorfall_growthmodel)

ggplot(indoor_height, aes(x = age, y = height_cm)) +
  geom_point(alpha = 1/10) +
  geom_smooth(method = "lm", col = "dodgerblue", se = FALSE) +
  ylab("Height (cm)") +
  xlab("Age (day)") +
  ggtitle("Indoor")

ggplot(filter(outdoor_height, season == "Summer 2016"), aes(x = age, y = height_cm)) +
  geom_point(alpha = 1/10) +
  geom_smooth(method = "lm", col = "dodgerblue", se = FALSE) +
  ylab("Height (cm)") +
  xlab("Age (day)") +
  ggtitle("Outdoor - Summer 2016")

ggplot(filter(outdoor_height, season == "Fall 2016"), aes(x = age, y = height_cm)) +
  geom_point(alpha = 1/10) +
  geom_smooth(method = "lm", col = "dodgerblue", se = FALSE) +
  ylab("Height (cm)") +
  xlab("Age (day)") +
  ggtitle("Outdoor - Fall 2016")

```

## Modeling - Sravani

```{r, echo = FALSE}
library(stringr)
joined_table <- traits %>%
  left_join(variables, by = 'variable_id') %>%
  left_join(cultivars, by = 'cultivar_id') %>%
  left_join(entities, by = 'entity_id') %>%
  left_join(sites, by = 'site_id') %>%
  left_join(treatments, by = 'treatment_id') %>%
  dplyr::select(trait_id, date, mean, variable_name, sitename, treatment_name, cultivar)

filtered_table <- filter(joined_table, variable_name %in% c("height", "canopy_cover", "canopy_height", "perimeter", "aboveground_dry_biomass", "leaf_length", "leaf_width", "plant_height", "aboveground_fresh_biomass", "growth_respiration_coefficient", "germination_score", "stem_diameter", "emergence_count")) %>%
  collect(n = Inf)

#Separating the filtered table into the outdoor and the indoor table
outdoortable = filtered_table %>%
  filter(str_detect(sitename, "Field Scanner"))

indoortable = filtered_table %>%
  filter(str_detect(sitename, "Danforth Plant Science Center")) 
```

###Outdoor

We first check if any of the variables in the outdoor table are missing values.

```{r}
sum(is.na(outdoortable$date))
sum(is.na(outdoortable$trait_id))
sum(is.na(outdoortable$mean))
sum(is.na(outdoortable$variable_name))
sum(is.na(outdoortable$treatment_name))
```

Since $treatment\_name$ is missing 697039 values, it is not considered in any of the modeling part.

As $date$ is missing 7020 values, we create a new dataset, $outdata$ that does not consider those rows which have missing date entries.

```{r}
a = which(is.na(outdoortable$date))
outdata = outdoortable[-a,]
```


```{r}
model1 = lm(log(mean + 1) ~ variable_name + trait_id + date, data = outdata)
summary(model1) 
summary(model1)$adj.r.squared
```

$trait\_id$ is insignificant while $variable_name$ and $date$ are. The $Adjusted R^2$ value is 0.9464145 indicating this model is pretty good.

Considering interaction terms, we get,

```{r}
model2 = lm(log(mean + 1) ~ variable_name * trait_id * date, data = outdata)
summary(model2) 
summary(model2)$adj.r.squared
```

This has a higher $Adjusted R^2$ value of 0.9593956.

```{r}
plot(model1, which = c(2))
plot(model2, which = c(2))
```

Both of the above models do not obey normal distribution.

```{r}
model3 = lm(log(mean + 1) ~ variable_name * date, data = outdata)
summary(model3) 
summary(model3)$adj.r.squared
```

```{r}
plot(model3, which = c(2))
```

###Indoor

```{r}
modeli1 = lm(mean ~ variable_name + trait_id + date + treatment_name, data = indoortable)
summary(modeli1)
```

```{r}
modeli2 = lm(mean ~ 1 + date * treatment_name, data = indoortable, subset = variable_name == 'perimeter')
plot(modeli2, which = c(2))
summary(modeli2)

modeli3 = lm(mean ~ 1 + date * treatment_name, data = indoortable, subset = variable_name =='plant_height')
plot(modeli3, which = c(2))

library(ggplot2) 
ggplot(indoortable, aes(date, mean, color = treatment_name))+
  geom_point() +
  geom_smooth(se = FALSE) +
  facet_wrap(~variable_name, scales = 'free')
```

```{r}
library(MASS)
library(faraway)
boxcox(modeli2, lambda = seq(-0.25, 0.75, by = 0.05), plotit = TRUE)
```

```{r}
modeli2box = lm((((mean ^ 0.25) - 1) / 0.25) ~  1 + date * treatment_name, data = indoortable, subset = variable_name == 'perimeter')
plot(modeli2box, which = c(2))
summary(modeli2box)
```

```{r}
library(MASS)
library(faraway)
boxcox(modeli3, lambda = seq(-0.25, 0.75, by = 0.05), plotit = TRUE)
```

```{r}
modeli3box = lm((((mean ^ 0.45) - 1) / 0.45) ~ 1 + date * treatment_name, data = indoortable, subset = variable_name =='plant_height')
plot(modeli3, which = c(2))
```

```{r}
total4 = which(cooks.distance(modeli2box) < 4 / length(cooks.distance(modeli2box)))
lev4 = which(hatvalues(modeli2box) > 2 * mean(hatvalues(modeli2box)))
indoordata4 = indoortable[-total4 -lev4,]
modeli2newbox = lm((((mean ^ 0.25) - 1) / 0.25) ~  1 + date * treatment_name, data = indoordata4, subset = variable_name == 'perimeter')
plot(modeli2newbox, which = c(2))
```

```{r}
plot(modeli2, which = c(2))
```

```{r}
plot(model2, which = c(1))
plot(model3, which = c(1))
plot(modeli2, which = c(1))
```

```{r}
#library(lmtest)
#bptest(model2)
#bptest(model3)
#bptest(modeli2)

```

For all of the above models, we see a small $p-value$, so we do reject the null of homoscedasticity and thus, the constant variance assumption is violated.

The $QQ plots$ for the above models violate the normality assumption.

So, we next try to remove the influential points.

#Remove influential points

```{r}
total = which(cooks.distance(model2) < 4 / length(cooks.distance(model2)))
lev = which(hatvalues(model2) > 2 * mean(hatvalues(model2)))
outtotdata = outdata[-total,]
model2new = lm(log(mean + 1) ~ variable_name * trait_id * date, data = outtotdata)
plot(model2new, which = c(2))
```

```{r}
total2 = which(cooks.distance(model3) < 4 / length(cooks.distance(model3)))
lev2 = which(hatvalues(model3) > 2 * mean(hatvalues(model3)))
outtotdata2 = outdata[-total2 - lev2,]
model3new = lm(mean ~ variable_name * date, data = outtotdata2, subset = cooks.distance(model3) < 4 / length(cooks.distance(model3)))
plot(model3new, which = c(2))
```

```{r}
total3 = which(cooks.distance(modeli2) < 4 / length(cooks.distance(modeli2)))
lev3 = which(hatvalues(modeli2) > 2 * mean(hatvalues(modeli2)))
indoordata2 = indoortable[-total3 - lev3,]
modeli2new = lm(log(mean + 1) ~ variable_name * date * treatment_name, data = indoordata2)
plot(modeli2new, which = c(2))
plot(modeli2, which = c(2))
```

# Modeling - Yuji

## Plots

```{r}
## indoor models
ggplot(indoor_height, aes(x = age, y = height_cm)) +
  geom_point(aes(color = treatment_name, alpha = 1/10)) +
  ylab("Plant Height (Indoor)") +
  facet_wrap(~treatment_name) +
  ggtitle("Indoor Plants by Treatment")

m_indoor_treat <- lm(height_cm ~ age*treatment_name, indoor_height)
summary(m_indoor_treat)

## no significant difference among treatment, except for a slight reduction for 40%

ggplot(indoor_height, aes(x = age, y = height_cm)) +
  geom_point(aes(color = cultivar, alpha = 1/10)) +
  ylab("Height (cm)") +
  facet_wrap(~cultivar) +
  ggtitle("Indoor Plants by Cultivar")

m_indoor_cultivar <- lm(height_cm ~ age*cultivar, indoor_height)
summary(m_indoor_cultivar)

## Growing rates: BTx642 = Tx430 > PI_564163 > TX7000

## outdoor model
ggplot(outdoor_height, aes(x = age, y = height_cm)) +
  geom_point(aes(color = season, alpha = 1/10)) +
  ylab("Height (cm)") +
  ggtitle("Outdoor Plants by Season")

m_out_season <- lm(height_cm ~ age*season, data = outdoor_height)
summary(m_out_season)

## growing faster in summer than in fall

```

## Put together

```{r}
whole_height <- rbind(indoor_height, outdoor_height)

ggplot(whole_height, aes(x = age, y = height_cm)) +
  geom_point(aes(color = indoor_outdoor, alpha = 1/10)) +
  ylab("Height (cm)") +
  ggtitle("Outdoor Plants by Season")


## Seems a good model: interaction term accounts for the different growth rates
m_whole <- lm(height_cm ~ age*indoor_outdoor, data = whole_height)
summary(m_whole)

## Significant difference between indoor and outdoor

```