---
title: "Importing & Exporting Tables"
author: "Ran Ji"
date: "5/30/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading BetyDB into environment and joining to make main table
```{r db-connection}
library(dplyr)

bety_src <- src_postgres(dbname = "bety", 
                password = 'bety', 
                host = 'terra-bety.default', 
                user = 'bety', 
                port = 5432)

original_traits <- tbl(bety_src, 'traits') %>%
  mutate (traits_id = id) 

original_variables <- tbl(bety_src, 'variables') %>%
  mutate(variable_id = id, variable_name = name)

traits <- tbl(bety_src, 'traits', n = Inf) %>%
  mutate (traits_id = id) %>%
  select (traits_id, site_id, specie_id, cultivar_id, date, mean, variable_id, method_id, treatment_id, entity_id)

variables <- tbl(bety_src, 'variables', n = Inf) %>%
  mutate(variable_id = id, variable_name = name) %>%
  select(variable_id, variable_name, description, units)

variables_local <- tbl(bety_src, 'variables', n = Inf) %>%
  mutate(variable_id = id, variable_name = name) %>%
  select(variable_id, variable_name, description, units) %>%
  collect()

cultivars <- tbl(bety_src, 'cultivars', n = Inf) %>%
  mutate(cultivar_id = id, cultivar = name) %>%
  select(cultivar_id, cultivar)

entities <- tbl(bety_src, 'entities', n = Inf) %>%
  mutate(entity_name = name, entity_id = id) %>%
  select(entity_name, entity_id)

sites <- tbl(bety_src, 'sites', n = Inf) %>%
  mutate(site_id = id) %>%
  select(site_id, city, state, country, notes, sitename)

treatments <- tbl(bety_src, 'treatments', n = Inf) %>%
  mutate(treatment_id = id, treatment_definition = definition, treatment_name = name) %>%
  select(treatment_id, treatment_name, treatment_definition) 

joined_table <- traits %>%
  left_join(variables, by = 'variable_id') %>%
  left_join(cultivars, by = 'cultivar_id') %>%
  left_join(entities, by = 'entity_id') %>%
  left_join(sites, by = 'site_id') %>%
  left_join(treatments, by = 'treatment_id') %>%
  select(traits_id, date, mean, variable_name, sitename, treatment_name, cultivar) %>%
  collect(n=Inf)

#sum(is.na(joined_table$cultivar))
#sum(!is.na(joined_table$cultivar))
```

## Taking subset of joined_table containing variables that we care about
```{r variable-tables}
filtered_table <- filter(joined_table, variable_name %in% c("height", "canopy_cover", "canopy_height", "perimeter", "aboveground_dry_biomass", "leaf_length", "leaf_width", "plant_height", "aboveground_fresh_biomass", "growth_respiration_coefficient", "germination_score", "stem_diameter", "emergence_count", "cultivar")) %>%
  collect(n = Inf)
```

## Indoor and outdoor tables
```{r}
indoor_table <- filtered_table[grepl("Danforth", filtered_table$sitename), ]
outdoor_table <- filtered_table[grepl("MAC", filtered_table$sitename), ]
```

## Compare Means
```{r}
mean(outdoor_table$mean[outdoor_table$variable_name == "height"])
mean(outdoor_table$mean[outdoor_table$variable_name == "canopy_height"])
mean(indoor_table$mean[indoor_table$variable_name == "plant_height"])
```

## Plots
```{r}
hist(outdoor_table$mean[outdoor_table$variable_name == "height"])
hist(indoor_table$mean[indoor_table$variable_name == "plant_height"]/100)
plot(indoor_table$date[indoor_table$variable_name == "plant_height"],indoor_table$mean[indoor_table$variable_name == "plant_height"])
plot(outdoor_table$date[outdoor_table$variable_name == "height"],outdoor_table$mean[outdoor_table$variable_name == "height"])
plot(outdoor_table$date[outdoor_table$variable_name == "canopy_height"],outdoor_table$mean[outdoor_table$variable_name == "canopy_height"])
```

## Summary to find dates
```{r}
summary(indoor_table$date[indoor_table$variable_name == "plant_height"])
summary(outdoor_table$date[outdoor_table$variable_name == "height"])
summary(outdoor_table$date[outdoor_table$variable_name == "canopy_height"])
```

## Filtering out 2017 data
```{r}
year_2017 <- joined_table[grepl(2017, joined_table$date), ]
unique(year_2017$variable_name)
```

## Filtering out and exploring data with nonempty cultivar
```{r}
cultivar_available <- joined_table[!is.na(joined_table$cultivar), ]

#unique(cultivar_available$sitename)
#unique(joined_table$sitename)

length(unique(cultivar_available$sitename))
length(unique(joined_table$sitename))

cultivar_sites <- unique(cultivar_available$sitename)
all_sites <- unique(joined_table$sitename)

#setdiff(all_sites, cultivar_sites)
```

## What are Season 1 and 2?
```{r}
range(joined_table$date[grepl('Season 1', joined_table$sitename) & !is.na(joined_table$date)])
range(joined_table$date[grepl('Season 2', joined_table$sitename)])
length(unique(joined_table$sitename[grepl('Season 1', joined_table$sitename)]))
length(unique(joined_table$date[grepl('Season 1', joined_table$sitename)]))
length(unique(joined_table$date[grepl('Season 1', joined_table$sitename)]))
```

##  Sorting by season and adding date columns
```{r}
library(lubridate)

indoor_table_plant_height <- indoor_table[indoor_table$variable_name == 'plant_height' & !is.na(indoor_table$date),]

##change units of plant_height from mm to cm
indoor_table_plant_height$mean = indoor_table_plant_height$mean *.1

indoor_start_date <- as.POSIXct("2014-05-27 12:00:00")
for (i in 1:length(indoor_table_plant_height$date)){
  indoor_table_plant_height$age[i] <- as.numeric(as.POSIXct(indoor_table_plant_height$date[i]) - indoor_start_date)
}

outdoor_table_canopy_height_summer <- outdoor_table[outdoor_table$variable_name == 'canopy_height' & outdoor_table$date %within% interval(as.POSIXct("2016-01-01 01:00:00"), as.POSIXct("2016-08-01 01:00:00")) & !is.na(outdoor_table$date), ]

outdoorsummer_start_date <- as.POSIXct("2016-04-19 12:00:00")
for (i in 1:length(outdoor_table_canopy_height_summer$date)){
  outdoor_table_canopy_height_summer$age[i] <- as.numeric(as.POSIXct(outdoor_table_canopy_height_summer$date[i]) - outdoorsummer_start_date)
}

outdoor_table_canopy_height_fall <- outdoor_table[outdoor_table$variable_name == 'canopy_height' & outdoor_table$date %within% interval(as.POSIXct("2016-08-01 01:00:00"), as.POSIXct("2016-12-01 01:00:00")) & !is.na(outdoor_table$date), ]

outdoorfall_start_date <- as.POSIXct("2016-08-03 12:00:00")
for (i in 1:length(outdoor_table_canopy_height_fall$date)){
  outdoor_table_canopy_height_fall$age[i] <- as.numeric(as.POSIXct(outdoor_table_canopy_height_fall$date[i]) - outdoorfall_start_date)
}

```

## Linear Regression!
```{r}
plot(x=outdoor_table_canopy_height_summer$date, y=outdoor_table_canopy_height_summer$mean, type = 'p')
plot(x=outdoor_table_canopy_height_fall$date, y=outdoor_table_canopy_height_fall$mean, type = 'p')

indoor_growthmodel <- lm(mean ~ age, data = indoor_table_plant_height)
summary(indoor_growthmodel)

outdoorsummer_growthmodel <- lm(mean ~ age, data = outdoor_table_canopy_height_summer)
summary(outdoorsummer_growthmodel)

outdoorfall_growthmodel <- lm(mean ~ age, data = outdoor_table_canopy_height_fall)
summary(outdoorfall_growthmodel)

plot(x=indoor_table_plant_height$age, y=indoor_table_plant_height$mean, type = 'p')  
abline(indoor_growthmodel)

plot(x=outdoor_table_canopy_height_summer$age, y=outdoor_table_canopy_height_summer$mean, type = 'p')  
abline(outdoorsummer_growthmodel)

plot(x=outdoor_table_canopy_height_fall$age, y=outdoor_table_canopy_height_fall$mean, type = 'p')  
abline(outdoorfall_growthmodel)


```



