---
title: "Modeling"
author: "Sravani Vadali"
date: "6/9/2017"
output: html_document
---

```{r, echo = FALSE}
library(stringr)

library(dplyr)

bety_src <- src_postgres(dbname = "bety", 
                password = 'bety', 
                host = 'terra-bety.default', 
                user = 'bety', 
                port = 5432)

# to see all available columns in traits table
original_traits <- tbl(bety_src, 'traits') %>%
  collect(n=1)

# local version of variables for reference
variables_local <- tbl(bety_src, 'variables', n = Inf) %>%
  mutate(variable_id = id, variable_name = name) %>%
  dplyr::select(variable_id, variable_name, description, units) %>%
  collect()

traits <- tbl(bety_src, 'traits', n = Inf) %>%
  mutate(trait_id = id) %>%
  dplyr::select(trait_id, site_id, specie_id, cultivar_id, date, mean, variable_id, method_id, treatment_id, entity_id)

variables <- tbl(bety_src, 'variables', n = Inf) %>%
  mutate(variable_id = id, variable_name = name) %>%
  dplyr::select(variable_id, variable_name)

cultivars <- tbl(bety_src, 'cultivars', n = Inf) %>%
  mutate(cultivar_id = id, cultivar = name) %>%
  dplyr::select(cultivar_id, cultivar)

entities <- tbl(bety_src, 'entities', n = Inf) %>%
  mutate(entity_name = name, entity_id = id) %>%
  dplyr::select(entity_name, entity_id)

sites <- tbl(bety_src, 'sites', n = Inf) %>%
  mutate(site_id = id) %>%
  dplyr::select(site_id, city, state, country, notes, sitename)

treatments <- tbl(bety_src, 'treatments', n = Inf) %>%
  mutate(treatment_id = id, treatment_definition = definition, treatment_name = name) %>%
  dplyr::select(treatment_id, treatment_name, treatment_definition) 

joined_table <- traits %>%
  left_join(variables, by = 'variable_id') %>%
  left_join(cultivars, by = 'cultivar_id') %>%
  left_join(entities, by = 'entity_id') %>%
  left_join(sites, by = 'site_id') %>%
  left_join(treatments, by = 'treatment_id') %>%
  dplyr::select(trait_id, date, mean, variable_name, sitename, treatment_name, cultivar)

filtered_table <- filter(joined_table, variable_name %in% c("height", "canopy_cover", "canopy_height", "perimeter", "aboveground_dry_biomass", "leaf_length", "leaf_width", "plant_height", "aboveground_fresh_biomass", "growth_respiration_coefficient", "germination_score", "stem_diameter", "emergence_count")) %>%
  collect(n = Inf)

#Separating the filtered table into the outdoor and the indoor table
outdoortable = filtered_table %>%
  filter(str_detect(sitename, "Field Scanner"))

indoortable = filtered_table %>%
  filter(str_detect(sitename, "Danforth Plant Science Center")) 
```

#Outdoor Analysis

We first check if any of the variables in the outdoor table are missing values.

```{r}
sum(is.na(outdoortable$date))
sum(is.na(outdoortable$trait_id))
sum(is.na(outdoortable$mean))
sum(is.na(outdoortable$variable_name))
sum(is.na(outdoortable$treatment_name))
```

Since $treatment\_name$ is missing 697039 values, it is not considered in any of the modeling part.

As $date$ is missing 7020 values, we create a new dataset, $outdata$ that does not consider those rows which have missing date entries.

```{r}
a = which(is.na(outdoortable$date))
outdata = outdoortable[-a,]
```


```{r}
model1 = lm(log(mean + 1) ~ variable_name + trait_id + date, data = outdata)
summary(model1) 
summary(model1)$adj.r.squared
```

$trait\_id$ is insignificant while $variable\_name$ and $date$ are. The $Adjusted R^2$ value is 0.9464145 indicating this model is pretty good.

Considering interaction terms, we get,

```{r}
model2 = lm(log(mean + 1) ~ variable_name * trait_id * date, data = outdata)
summary(model2) 
summary(model2)$adj.r.squared
```

This has a higher $Adjusted R^2$ value of 0.9593952.

```{r}
model1b = lm(log((mean + 1)^2) ~ variable_name + trait_id + date, data = outdata)
model2b = lm(log((mean + 1)^2) ~ variable_name * trait_id * date, data = outdata)
summary(model1b)$adj.r.squared
summary(model2b)$adj.r.squared
```


```{r}
plot(model1, which = c(2))
```

```{r}
plot(model2, which = c(2))
```

```{r}
plot(model1b, which = c(2))
```

```{r}
plot(model2b, which = c(2))
```

All of the above models do not obey normal distribution.

```{r}
model3 = lm(log(mean + 1) ~ variable_name * date, data = outdata)
summary(model3) 
summary(model3)$adj.r.squared
```

```{r}
plot(model3, which = c(2))
```

So, we next try to remove the influential points.

##Remove influential points


```{r}
total = which(cooks.distance(model2) < 4 / length(cooks.distance(model2)))
lev = which(hatvalues(model2) > 2 * mean(hatvalues(model2)))
outtotdata = outdata[-total,]
model2new = lm(log(mean + 1) ~ variable_name * trait_id * date, data = outtotdata)
plot(model2new, which = c(2))
```

```{r}
summary(model2new)$adj.r.squared
```


The $model2new$ is the best model when we consider the Outdoor table. This is because it has a good $Adjusted\;R^2$ value of 0.8936986 and the QQ plot indicates that it almost follows a normal distribution.


#Indoor Analysis

```{r}
modeli1 = lm(mean ~ variable_name + trait_id + date + treatment_name, data = indoortable)
summary(modeli1)
```

```{r}
modeli2 = lm(mean ~ 1 + date * treatment_name, data = indoortable, subset = variable_name == 'perimeter')
plot(modeli2, which = c(2))
summary(modeli2)


modeli3 = lm(mean ~ 1 + date * treatment_name, data = indoortable, subset = variable_name =='plant_height')
plot(modeli3, which = c(2))

library(ggplot2) 
ggplot(indoortable, aes(date, mean, color = treatment_name))+
  geom_point() +
  geom_smooth(se = FALSE) +
  facet_wrap(~variable_name, scales = 'free')
```


```{r}
modeli2box = lm((((mean ^ 0.15) - 1) / 0.15) ~  1 + date * treatment_name, data = indoortable, subset = variable_name == 'perimeter')
plot(modeli2box, which = c(2))
summary(modeli2box)
```


```{r}
modeli3box = lm((((mean ^ 0.45) - 1) / 0.45) ~ 1 + date * treatment_name, data = indoortable, subset = variable_name =='plant_height')
plot(modeli3, which = c(2))
```

```{r}
total4 = which(cooks.distance(modeli2box) < 4 / length(cooks.distance(modeli2box)))
lev4 = which(hatvalues(modeli2box) > 2 * mean(hatvalues(modeli2box)))
indoordata4 = indoortable[-total4 -lev4,]
modeli2newbox = lm((((mean ^ 0.15) - 1) / 0.15) ~  1 + date * treatment_name, data = indoordata4, subset = variable_name == 'perimeter')
plot(modeli2newbox, which = c(2))
```

```{r}
summary(modeli2newbox)$adj.r.squared
```


The $modeli2newbox$ is the best model when we consider the Outdoor table. This is because it has a good $Adjusted\;R^2$ value of 0.7636435 and the QQ plot indicates that it almost follows a normal distribution.

