---
title: "Importing & Exporting Tables"
author: "Sravani Vadali"
date: "5/30/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading BetyDB into environment and joining to make main table
```{r db-connection}
library(dplyr)

bety_src <- src_postgres(dbname = "bety", 
                password = 'bety', 
                host = 'terra-bety.default', 
                user = 'bety', 
                port = 5432)

original_traits <- tbl(bety_src, 'traits') %>%
  mutate (traits_id = id)

original_variables <- tbl(bety_src, 'variables') %>%
  mutate(variable_id = id, variable_name = name)

traits <- tbl(bety_src, 'traits', n = Inf) %>%
  mutate (traits_id = id) %>%
  select (traits_id, site_id, specie_id, cultivar_id, date, mean, variable_id, method_id, treatment_id, entity_id)

variables <- tbl(bety_src, 'variables', n = Inf) %>%
  mutate(variable_id = id, variable_name = name) %>%
  select(variable_id, variable_name, description, units)

variables_local <- tbl(bety_src, 'variables', n = Inf) %>%
  mutate(variable_id = id, variable_name = name) %>%
  select(variable_id, variable_name, description, units) %>%
  collect()

cultivars <- tbl(bety_src, 'cultivars', n = Inf) %>%
  mutate(cultivar_id = id, cultivar = name) %>%
  select(cultivar_id, cultivar)

entities <- tbl(bety_src, 'entities', n = Inf) %>%
  mutate(entity_name = name, entity_id = id) %>%
  select(entity_name, entity_id)

sites <- tbl(bety_src, 'sites', n = Inf) %>%
  mutate(site_id = id) %>%
  select(site_id, city, state, country, notes, sitename)

treatments <- tbl(bety_src, 'treatments', n = Inf) %>%
  mutate(treatment_id = id, treatment_definition = definition, treatment_name = name) %>%
  select(treatment_id, treatment_name, treatment_definition) 

joined_table <- traits %>%
  left_join(variables, by = 'variable_id') %>%
  left_join(cultivars, by = 'cultivar_id') %>%
  left_join(entities, by = 'entity_id') %>%
  left_join(sites, by = 'site_id') %>%
  left_join(treatments, by = 'treatment_id') %>%
  select(traits_id, date, mean, variable_name, sitename, treatment_name)
```

##Taking subset of joined_table containing variables that we care about
```{r variable-tables}
filtered_table <- filter(joined_table, variable_name %in% c("height", "canopy_cover", "canopy_height", "perimeter", "aboveground_dry_biomass", "leaf_length", "leaf_width", "plant_height", "aboveground_fresh_biomass", "growth_respiration_coefficient", "germination_score", "stem_diameter", "emergence_count")) %>%
  collect(n = Inf)
```

```{r}
length(which(filtered_table$variable_name == 'height'))
```

```{r}
length(which(filtered_table$variable_name == 'aboveground_dry_biomass'))
```

```{r}
library(stringr)

outdoor_height = filtered_table %>%
  filter(str_detect(sitename, "Field Scanner")) %>%
  filter(str_detect(variable_name, "height"))

indoor_height = filtered_table %>%
  filter(str_detect(sitename, "Danforth Plant Science Center")) %>%
  filter(str_detect(variable_name, "height"))

```

```{r}
#Separating the filtered table into the outdoor and the indoor table
outdoortable = filtered_table %>%
  filter(str_detect(sitename, "Field Scanner"))

indoortable = filtered_table %>%
  filter(str_detect(sitename, "Danforth Plant Science Center")) 
```


###Outdoor

We first check if any of the variables in the outdoor table are missing values.

```{r}
sum(is.na(outdoortable$date))
sum(is.na(outdoortable$traits_id))
sum(is.na(outdoortable$mean))
sum(is.na(outdoortable$variable_name))
sum(is.na(outdoortable$treatment_name))
```

Since $treatment\_name$ is missing 697039 values, it is not considered in any of the modeling part.

As $date$ is missing 7020 values, we create a new dataset, $outdata$ that does not consider those rows which have missing date entries.

```{r}
a = which(is.na(outdoortable$date))
outdata = outdoortable[-a,]
```


```{r}
model1 = lm(log(mean + 1) ~ variable_name + traits_id + date, data = outdata)
summary(model1) 
summary(model1)$adj.r.squared
```

$traits\_id$ is insignificant while $variable_name$ and $date$ are. The $Adjusted R^2$ value is 0.8082532 indicating this model is pretty good.

Considering interaction terms, we get,

```{r}
model2 = lm(log(mean + 1) ~ variable_name * traits_id * date, data = outdata)
summary(model2) 
summary(model2)$adj.r.squared
```

This has a higher $Adjusted R^2$ value of 0.9529574.

```{r}
plot(model1, which = c(2))
plot(model2, which = c(2))
```

Both of the above models do not obey normal distribution.

```{r}
model3 = lm(log(mean + 1) ~ variable_name * date, data = outdata)
summary(model3) 
summary(model3)$adj.r.squared
```

```{r}
plot(model3, which = c(2))
```

###Indoor

```{r}
modeli1 = lm(mean ~ variable_name + traits_id + date + treatment_name, data = indoortable)
summary(modeli1)
```

```{r}
modeli2 = lm(log(mean + 1) ~ variable_name * date * treatment_name, data = indoortable)
summary(modeli2)
```

```{r}
plot(modeli2, which = c(2))
```

```{r}
plot(model2, which = c(1))
plot(model3, which = c(1))
plot(modeli2, which = c(1))
```

```{r}
library(lmtest)
bptest(model2)
bptest(model3)
bptest(modeli2)

```

For all of the above models, we see a small $p-value$, so we do reject the null of homoscedasticity and thus, the constant variance assumption is violated.

The $QQ plots$ for the above models violate the normality assumption.

So, we next try to remove the influential points.

#Remove influential points

```{r}
total = which(cooks.distance(model2) < 4 / length(cooks.distance(model2)))
lev = which(hatvalues(model2) > 2 * mean(hatvalues(model2)))
outtotdata = outdata[-total,]
model2new = lm(log(mean + 1) ~ variable_name * traits_id * date, data = outtotdata)
plot(model2new, which = c(2))
```

```{r}
total2 = which(cooks.distance(model3) < 4 / length(cooks.distance(model3)))
lev2 = which(hatvalues(model3) > 2 * mean(hatvalues(model3)))
outtotdata2 = outdata[-total2 - lev2,]
model3new = lm(mean ~ variable_name * date, data = outtotdata2, subset = cooks.distance(model3) < 4 / length(cooks.distance(model3)))
plot(model3new, which = c(2))
```

```{r}
total3 = which(cooks.distance(modeli2) < 4 / length(cooks.distance(modeli2)))
lev3 = which(hatvalues(modeli2) > 2 * mean(hatvalues(modeli2)))
indoordata2 = indoortable[-total3 - lev3,]
modeli2new = lm(log(mean + 1) ~ variable_name * date * treatment_name, data = indoordata2)
plot(modeli2new, which = c(2))
```


###Both

```{r}
#modelb1 = lm(mean ~., data = filtered_table) #Not great values anyway. Don't use
#summary(modelb1)
```


```{r}
plot(outdoor_height$date, outdoor_height$mean)
```




